Items

### 2021-10-25 proposal.Rmd
changes to theory

abstract          : "Privacy is an important societal topic. Living in an information age, people constantly have to decide what information to share, which service to use, or when to communicate. All of these decisions are reflective of and determined by the users' need for privacy. It is relevant to understand who needs more and who needs less privacy, because desiring privacy often requires justification. For example, the nothing-to-hide statement -- i.e., 'someone who has nothing to hide has nothing to fear' -- imply that people who desire privacy are suspicious. Although such suspicions might be justified in some cases, there exist many alternative legitimate explanations why people desire privacy. For example, they could be more introverted, hesitant, creative, or prudent. In this study we, hence, plan to explore the relation between personality and the need for privacy. Personality factors and facets will be operationalized using the HEXACO personality inventory. Need for privacy will be captured with a multidimensional approach, including informational and social privacy, need for privacy from government agencies, or need for privacy from companies. We will collect a sample of more than 800 respondents, which are representative of the US in terms of age, gender, and ethnicity. The relations between personality and privacy will be explored using structural equation modeling."
keywords          : "Privacy, need for privacy, personality, HEXACO, structural equation modeling"

@@

Amidst the increasing digitization of everyday life, 
<!-- which has led to several sweeping societal changes such as the commodification and monetization of personal information [@sevignaniPrivacyCapitalismAge2016],  -->
privacy has become a major topic of public and academic interest. 
Despite the topic's importance, to date we still know surprisingly little about the relation between privacy and personality [@masurSituationalPrivacySelfdisclosure2018, p. 155]. 
Why do some people feel they need or desire more privacy than others, and how do these people differ from one another?

We believe it is relevant to understand better this research question, because people who desire privacy are often asked to justify their need for privacy. 
For example, well-known statements such as 'who has nothing to hide has nothing to fear' imply that people who desire privacy are suspicious.
Granted, it is only logical that people who commit crimes and who are insincere in fact benefit from more privacy.
However, there exist many other alternative reasons why people need privacy.
For example, it could also be that people who need more privacy are just more introverted, hesitant, creative, or prudent. 
We therefore believe that a better understanding of the relation between personality and privacy is relevant.

This research question is relevant also from an academic perspective. 
Several theories argue that personality determines privacy behaviors [@masurSituationalPrivacySelfdisclosure2018, p. 155]. 
However, to date there is almost no empirical research that can be used to develop well-informed hypotheses. 
As a result, with this paper we would like to answer the following question:  
What personality factors and facets best explain peoples' felt need for privacy?

@@

The theoretical concept of privacy is complicated and contested [@nissenbaumPrivacyContextTechnology2010, p. 71].
Thus, we first outline our own understanding of privacy. 
First and foremost, privacy captures a _withdrawal_ from others or from society in general [@westinPrivacyFreedom1967].
This withdrawal (b) happens _voluntarily_ and is under a person's _control_ [@westinPrivacyFreedom1967]. 
Several models suggest that privacy is multi-dimensional. 
For example, in a theory-driven treatise @burgoonPrivacyCommunication1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. 
@pedersenDimensionsPrivacy1979, by contrast, conducted an empirical factor analysis of overall 94 items and found six dimensions of privacy: 
reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. 
@schwartzSocialPsychologyPrivacy1968 or @masurPrivatheitOnlineKommunikation2018 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies or business companies).

@@

<!-- In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually feel they need privacy.  -->
According to @trepteNeedPrivacy2017, the need for privacy is a secondary need---it is not an end in itself, but rather a way to satisfy other more fundamental needs such as safety, sexuality, recovery, or contemplation.
Specifically, @westinPrivacyFreedom1967 defined four ultimate purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (the desire to avoid being manipulated and dominated), (3) emotional release (the release of tension from social role demands), and (4) protected communication (the ability to foster intimate relationships). 
Not least, privacy facilitates self-disclosure [@dienlinPrivacyProcessModel2014], which is necessary for attaining social support, initiating relationships, and getting close to other people [@omarzuDisclosureDecisionModel2000].

On the other hand, however, privacy can also have negative aspects. 
It is possible to have too much privacy.
Human beings are inherently social, and being overly cut-off from others can diminish flourishing, nurture deviant behavior, or introduce power asymmetries [@altmanEnvironmentSocialBehavior1975]. 
Privacy fostering self-disclosure also presents a potential risk, because others might disagree, disapprove, or misuse the information in other contexts [@petronioCommunicationPrivacyManagement2010]. 
Privacy can also help conceal power-asymmetries, wrongdoing, or crimes such as violence or theft.
The dialectical tension between the positive and negative aspects of privacy thus causes variability across individuals in their need for privacy.

@@

So far, not a lot of studies have analyzed the relation between personality and need for privacy explicitly.
We are aware of only two studies that conducted an empirical analysis [@pedersenPersonalityCorrelatesPrivacy1982; @hosmanRelationshipsNeedPrivacy1991].
And as there is no established theory connecting privacy and personality, it is difficult to formulate precise and well-informed hypotheses. 
As a result, in this study we adopt a more _exploratory_ perspective.
We implement a large-scale perspective on personality, in order not to miss potentially relevant personality factors and facets.

To this end, we build on the HEXACO inventory of personality [@leePsychometricPropertiesHEXACO1002018].
The HEXACO model stands in the tradition of Big Five approach [@johnBigFiveTrait1999] and represents a broad understanding of personality.
It measures overall six factors (see below), which have four specific facets each. 
We build not only on general factors but also specific facets because we do not expect that the even more specific need for privacy dimensions will relate closely to overarching personality factors.
(For example, privacy concerns, a variable conceptually close to need for privacy, shows only small relations to the Big Five factors; @bansalImpactPersonalDispositions2010; @junglasPersonalityTraitsConcern2008.)
Another reason for choosing the HEXACO model was that in addition to the Big Five factors the HEXACO model includes a sixth one labeled Honesty Humility, 
<!-- plus another facet labeled Altruism,  -->
which seems promising to investigate the nothing to hide argument.
<!-- Hence, in order to increase precision we follow Paunonen's [-@paunonenBigFiveFactors2001] advice and also include specific personality facets. -->
<!-- To illustrate, instead of dwelling on the general factor of _Extraversion_, we instead focused on the facet _Gregariousness_, which we consider to be more pertinent.  -->
In what follows, we briefly present all factors and provide some tentative thoughts on how they and several selected facets might relate to privacy.

@@

Honesty-Humility consists of the facets sincerity, fairness, greed avoidance, and modesty.
We could imagine that honesty-humility relates to need for privacy for several reasons.
First, one can often hear that people need privacy because they have something to hide (are hence less honest, sincere, or fair).
The so-called _nothing-to-hide argument_ states that “If you have nothing to hide, you have nothing to fear.” 
<!-- As described by Solove, the nothing-to-hide argument says that -->
The argument for example implies that data mining and surveillance by government entities “is not likely to be threatening to the privacy of law-abiding citizens. 
Only those who are engaged in illegal activities have a reason to hide this information” [@soloveVeGotNothing2007, p. 753]. 
<!-- Hence, a potential predictor of why people need privacy could be how honest, correct, sincere, or benign they are. -->

<!-- Because integrity is a delicate concept, let us first try to define it conceptually.  -->
<!-- Although in terms of a scientific definition there is no consensus, most scholars seem to agree that integrity “incorporates a tendency to comply with social norms, avoid deviant behavior, and embrace a sense of justice, truthfulness, and fairness” [@Connelly.2006, p. 82].  -->
<!-- In order to sidestep the (very legitimate) philosophical debates about what constitutes integrity and what not, we hence follow @Paunonen.2002 and adopt a lowest common denominator definition, which means that we only consider participating in explicitly socially-sanctioned or illegal activities as a sign of lack of integrity. -->
<!-- It is possible to think of theoretical arguments for why lack of integrity might correlate positively with a person's felt need for privacy.  -->
Logically, people who actually commit crimes may face even greater risk from self-disclosure compared to others, because government agencies and people would sanction their activities [@petronioCommunicationPrivacyManagement2010]. 
Hence, the government and other people are more likely to be perceived as a threat, which should render anonymity a resource. 
As a consequence, people with lower honesty and humility might desire more privacy as a means to mitigate their felt risk [@altmanPrivacyConceptualAnalysis1976].
<!-- There are also a few empirical studies that imply---at least indirectly---a relation between privacy and integrity.  -->

Empirical studies have found that surveillance can indeed reduce cheating behaviors [@corcoranMoralityconscienceGuiltScale1987; @coveySelfmonitoringSurveillanceIncentive1989]. 
@coveySelfmonitoringSurveillanceIncentive1989 for example asked students to solve an impossible maze. 
In the high surveillance condition, the experimenter stood in front of the students and closely monitored their behavior. 
In the low surveillance condition, the experimenter remained behind the students where he or she could not see the students. 
Results showed greater cheating among students in the low surveillance condition, suggesting that in situations with less privacy, people show more honesty (i.e., fewer cheating behaviors). 
Next, in a longitudinal sample with 457 respondents in Germany [@treptePrivacySelfdisclosureSocial2013], people who felt they needed more privacy were also less authentic (and therefore, arguably, also less honest and sincere) on their online social network profiles (_r_ = -.48) and less authentic in their personal relationships (_r_ = -.28). 
<!-- Given the argument that authenticity is a subset of integrity [@Sheldon.2004], one could hence also reason that the concept of integrity might relate to a person’s perceived need for privacy.  -->

In conclusion, we could imagine that lack of honesty may indeed relate to an increased felt need for privacy, especially when it comes to government surveillance.
<!-- , as governments have the legitimate power to prosecute illegal activities.  -->
<!-- Next, it seems plausible that lack of integrity may relate to an increased need for anonymity, as anonymity makes it more difficult for both government and social agents to identify and address potential wrongdoers.  -->
<!-- Finally, lack of integrity may also relate to an increased need for privacy from other people, as most other people will disapprove of immoral or illegal activities, and could even reveal those activities to authorities. -->

>>>>

abstract          : "Privacy is an important societal topic. Living in an information age, people constantly have to decide what information to share, which service to use, when to communicate. All of these decisions are reflective of and determined by the users' need for privacy. We believe that it is relevant to understand who needs more and who needs less privacy, for example because desiring privacy often requires justification. The nothing-to-hide statement -- i.e., 'someone who has nothing to hide has nothing to fear' -- implies that people who desire privacy are suspicious. Although such suspicions might be justified in some cases, there are many alternative legitimate explanations as to why people desire privacy. For example, they could be more introverted, hesitant, creative, or prudent. In this study we hence plan to explore the relation between personality and the need for privacy. Personality factors and facets will be operationalized using the HEXACO personality inventory. Need for privacy will be captured with a multidimensional approach, including informational and social privacy, need for privacy from government agencies, or need for privacy from companies. We will collect a sample of more than 800 respondents, which will be representative of the US in terms of age, gender, and ethnicity. The relations between personality and privacy will be explored using structural equation modeling."
keywords          : "Privacy, need for privacy, personality, HEXACO, structural equation modeling"

@@

Amidst the increasing digitization of everyday life, 
<!-- which has led to several sweeping societal changes such as the commodification and monetization of personal information [@sevignaniPrivacyCapitalismAge2016],  -->
privacy has become a major topic of public and academic interest. 
Despite the topic's importance, to date we still know surprisingly little about the relation between privacy and personality [@masurSituationalPrivacySelfdisclosure2018, p. 155]. 
Why do some people feel they need more privacy than others, and how do these people differ from one another?

We believe it is relevant to understand better this research question, because people who desire privacy are often asked to justify themselves. 
For example, the so-called _nothing-to-hide argument_ states that “If you have nothing to hide, you have nothing to fear”. 
<!-- As described by Solove, the nothing-to-hide argument says that -->
The argument implies that people who desire privacy are suspicious. 
For example, once can often hear that data mining and surveillance by government entities “is not likely to be threatening to the privacy of law-abiding citizens. 
Only those who are engaged in illegal activities have a reason to hide this information” [@soloveVeGotNothing2007, p. 753].
And granted, it is only logical that people who commit crimes and who are insincere would in fact benefit from more privacy.
However, there exist many other alternative reasons as to why people need privacy.
For example, it could also be that people who need more privacy are just more introverted, hesitant, creative, or prudent. 
We therefore believe that a better understanding of the relation between personality and privacy is relevant from a societal perspective.

Also from an academic perspective, this research question is topical. 
Several theories argue that personality determines privacy behaviors [@masurSituationalPrivacySelfdisclosure2018, p. 155]. 
However, to date there is almost no empirical research that can be used to develop well-informed hypotheses. 
As a result, with this paper we would like to answer the following question:
What personality factors and facets best explain peoples' felt need for privacy?

@@

The theoretical concept of privacy is complicated and contested [@nissenbaumPrivacyContextTechnology2010, p. 71].
Thus, we first outline our own understanding of privacy. 
First and foremost, privacy captures a _withdrawal_ from others, or from society in general [@westinPrivacyFreedom1967].
This withdrawal (b) happens _voluntarily_ and is under a person's _control_ [@westinPrivacyFreedom1967]. 
Several models suggest that privacy is multi-dimensional. 
For example, in a theory-driven treatise @burgoonPrivacyCommunication1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. 
@pedersenDimensionsPrivacy1979 conducted an empirical factor analysis of overall 94 items and found six dimensions of privacy: 
reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. 
@schwartzSocialPsychologyPrivacy1968 and @masurPrivatheitOnlineKommunikation2018 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies or business companies).

@@

<!-- In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually feel they need privacy.  -->
According to @trepteNeedPrivacy2017, the need for privacy is a secondary need---it is not an end in itself, but rather a means to satisfy other more fundamental needs such as safety, sexuality, recovery, or contemplation.
Specifically, @westinPrivacyFreedom1967 defined four ultimate purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (the desire to avoid being manipulated and dominated), (3) emotional release (the release of tension from social role demands), and (4) protected communication (the ability to foster intimate relationships). 
Not least, privacy facilitates self-disclosure [@dienlinPrivacyProcessModel2014], which is necessary for attaining social support, initiating relationships, and getting close to other people [@omarzuDisclosureDecisionModel2000].

On the other hand, however, privacy can also have negative aspects. 
It is possible to have too much privacy.
Human beings are inherently social, and being overly cut-off from others can diminish flourishing, nurture deviant behavior, or introduce power asymmetries [@altmanEnvironmentSocialBehavior1975]. 
The fact that privacy fosters self-disclosure also presents a potential risk, because others might disagree, disapprove, or misuse the information in other contexts [@petronioCommunicationPrivacyManagement2010]. 
Privacy can also help conceal power-asymmetries, wrongdoing, or crimes such as violence or theft.
The dialectical tension between the positive and negative aspects of privacy thus might cause variability across individuals in their need for privacy.

@@

So far, not a lot of studies have analyzed the relation between personality and need for privacy explicitly.
We are aware of only two studies that conducted an empirical analysis [@pedersenPersonalityCorrelatesPrivacy1982; @hosmanRelationshipsNeedPrivacy1991].
And as there is no established theory connecting privacy and personality, it is difficult to formulate precise and well-informed a priori hypotheses. 
As a result, in this study we adopt a more _exploratory_ perspective.
We implement a large-scale operationalization on personality, in order not to miss potentially relevant personality factors and facets.

To this end, we build on the HEXACO inventory of personality [@leePsychometricPropertiesHEXACO1002018].
The HEXACO model stands in the tradition of the Big Five approach [@johnBigFiveTrait1999]
<!-- and represents a broad understanding of personality-->
.
It measures overall six factors (see below), which have four specific facets each. 
We include also the specific facets because we do not expect that the even more specific need for privacy dimensions will relate closely to overarching general personality factors.
(For example, privacy concerns, a variable conceptually close to need for privacy, shows only small relations to the Big Five factors; @bansalImpactPersonalDispositions2010; @junglasPersonalityTraitsConcern2008.)

Another reason for choosing the HEXACO model was that in addition to the Big Five factors the HEXACO model includes a sixth one labeled Honesty Humility, 
<!-- plus another facet labeled Altruism,  -->
which seems promising to investigate the nothing-to-hide-argument.
<!-- Hence, in order to increase precision we follow Paunonen's [-@paunonenBigFiveFactors2001] advice and also include specific personality facets. -->
<!-- To illustrate, instead of dwelling on the general factor of _Extraversion_, we instead focused on the facet _Gregariousness_, which we consider to be more pertinent.  -->
In what follows, we briefly present all factors and provide some tentative thoughts on how they and several selected facets might relate to privacy.

@@

Honesty-Humility consists of the facets sincerity, fairness, greed avoidance, and modesty.
According to the nothing-to-hide argument, one could assume that people might need privacy because they have something to hide---that is, because they are hence less honest, sincere, or fair.
<!-- Hence, a potential predictor of why people need privacy could be how honest, correct, sincere, or benign they are. -->
<!-- Because integrity is a delicate concept, let us first try to define it conceptually.  -->
<!-- Although in terms of a scientific definition there is no consensus, most scholars seem to agree that integrity “incorporates a tendency to comply with social norms, avoid deviant behavior, and embrace a sense of justice, truthfulness, and fairness” [@Connelly.2006, p. 82].  -->
<!-- In order to sidestep the (very legitimate) philosophical debates about what constitutes integrity and what not, we hence follow @Paunonen.2002 and adopt a lowest common denominator definition, which means that we only consider participating in explicitly socially-sanctioned or illegal activities as a sign of lack of integrity. -->
<!-- It is possible to think of theoretical arguments for why lack of integrity might correlate positively with a person's felt need for privacy.  -->
Logically, people who actually commit crimes may face even greater risk from self-disclosure compared to others, because government agencies and people would sanction their activities [@petronioCommunicationPrivacyManagement2010]. 
Hence, the government and other people are more likely to be perceived as a threat, which should render privacy a resource. 
As a consequence, once could argue that people with lower honesty and humility might desire more privacy as a means to mitigate their felt risk [@altmanPrivacyConceptualAnalysis1976].
<!-- There are also a few empirical studies that imply---at least indirectly---a relation between privacy and integrity.  -->

Empirical studies have found that surveillance can indeed reduce cheating behaviors [@corcoranMoralityconscienceGuiltScale1987; @coveySelfmonitoringSurveillanceIncentive1989]. 
@coveySelfmonitoringSurveillanceIncentive1989 for example asked students to solve an impossible maze. 
In the high surveillance condition, the experimenter stood in front of the students and closely monitored their behavior. 
In the low surveillance condition, the experimenter remained behind the students where they could not see the students. 
Results showed greater cheating among students in the low surveillance condition, suggesting that in situations with less privacy, people show more honesty (i.e., fewer cheating behaviors). 
Next, in a longitudinal sample with 457 respondents in Germany [@treptePrivacySelfdisclosureSocial2013], people who felt they needed more privacy were also less authentic (and therefore, arguably, also less honest and sincere) on their online social network profiles (_r_ = -.48) and less authentic in their personal relationships (_r_ = -.28). 
<!-- Given the argument that authenticity is a subset of integrity [@Sheldon.2004], one could hence also reason that the concept of integrity might relate to a person’s perceived need for privacy.  -->

In conclusion, it seems possible that lack of honesty may indeed relate to an increased felt need for privacy, especially when it comes to government surveillance.
<!-- , as governments have the legitimate power to prosecute illegal activities.  -->
<!-- Next, it seems plausible that lack of integrity may relate to an increased need for anonymity, as anonymity makes it more difficult for both government and social agents to identify and address potential wrongdoers.  -->
<!-- Finally, lack of integrity may also relate to an increased need for privacy from other people, as most other people will disapprove of immoral or illegal activities, and could even reveal those activities to authorities. -->
